---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
require(AGOtools)
require(tidyverse)
require(quickpsy)
```

# AGOtools

<!-- badges: start -->
<!-- badges: end -->

AGOtools is an R package providing some useful tools for the analysis of empirical data (plot themes for publication-ready plots using R's ggplot, outlier detection using a Tukey criterion, helper functions for fitting psychometric functions with the R package quickpsy (https://github.com/danilinares/quickpsy) etc.).

## Project Status

This Project is currently under construction. 

## Installation

Install AGOtools with:

```
devtools::install_github("AGOberfeld/AGOtools")
```

## Dependencies

If you want to use the helper fuznctions for quickpsy, make sure the correct Quickpsy-version (dev-version from Github, https://github.com/danilinares/quickpsy) is installed:

```
devtools::install_github("danilinares/quickpsy")
```


If you have already installed the official version from CRAN, remove quickpsy before installing it again from github.

```
remove.packages("quickpsy")

# install.packages("devtools")
require(devtools)
devtools::install_github("danilinares/quickpsy")

```

## Functions

### set_options

Sets the color palette of ggplots to default values. By default the matplotlib colors are chosen. Also sets the number format for the R output and loads some default packages (e.g., `tidyverse`, `pracma`, `extrafont`).

```{r, echo =TRUE}
set_options()
```


### tukey

Adds variables to an input dataset (data) which indicate if observations on a specified variable (dv) are outliers according to a "nonparametric"  criterion proposed by John Tukey. Values of variable dv more than k`*`IQR (interquartile range) below the first quartile (25% quantile) or more than k`*`IQR above the third quartile (75% quantile) are flagged as outliers.
The argument tukey_crit can be used to modify the factor k (default: tukey_crit = 3).

Identify outliers without excluding them:
```{r, echo =TRUE}
tukey(data = tukey_data , 
      dv = x, 
      tukey_crit=3, 
      exclude = FALSE)

tukey_data  %>% 
  tukey(x, exclude = FALSE) %>%
  ggplot(aes(y = x,x = ""))+
  geom_boxplot(width=0.3,
               outlier.shape = NA)+
  geom_jitter(aes(color = factor(x_outlierTukey)), 
              position = position_jitter(0.15), 
              alpha = 0.4)
```

Identify outliers and exclude them:
```{r, echo =TRUE}
tukey_data  %>%
	tukey(x, exclude = TRUE)
```

Outlier exclusion groupwise:
```{r, echo =TRUE}
tukey_data  %>% 
  group_by(group) %>% 
  tukey(x)

tukey_data  %>% 
  filter(group %in% c("A","B","C")) %>% 
  group_by(group) %>% 
  tukey(x, 
	      exclude = FALSE) %>%
  ggplot()+
  geom_jitter(aes(x = group, y = x, color = factor(x_outlierTukey)), alpha = 0.4, position = position_jitter(0.25))+
  plotThemeAGO()
```


 
 

Returns a list of variables and adds them to the initial data set:

`trialsInSet` = number of trials in the subset of trials for which the outlier detection is run  
`IQR` = inter-quartile range  
`Quant25` = 25% quantile  
`Quant75` = 75% quantile  
`outlierTukeyLow` = indicates whether the value of dv on a given trial is lower than Tukey_lower_limit (25% quantile minus tukey_crit* IQR)  (1) or not (0)  
`outlierTukeyHigh` = indicates whether the value of dv on a given trial is higher than Tukey_upper_limit (75% quantile plus tukey_crit* IQR)  (1) or not (0)  
`outlierTukey` = indicates if dv for a given trial exceeds the lower or the higher criterion (1) or is within both criteria (0)  
`Tukey_lower_limit` = highest value above which values are excluded  
`Tukey_upper_limit` = lowest value below which values are excluded  

### tidyQuickPsy

This function is used to facilitate the analysis of psychometric functions fitted with quickpsy (https://github.com/danilinares/quickpsy).

Takes a quickpsy-object and turns it as a tibble.

Use the following arguments in the quickpsy function:   
`d` = data  
`x` = Name of the explanatory variable (e.g. TTC)  
`k` = Name of the variable containing the number of "positive" decisions per group (= unique combination of the grouping variables) 
`n` = Name of the variable containing the number of trials per group (= unique combination of the grouping variables) 
`grouping` = concatenated vector of the variables that define the groups of trials (i.e., rows in the input dataset) to which separate psychometric functions are fit (e.g., a vector containing the participant code variable and the variables defining separate experimental conditions) 


```{r, warning=FALSE, echo =TRUE}
data <- streetcrossing %>% 
  filter(vp_code == "vp0001")

qp <- quickpsy(d = data,  
                x = track_TTC, 
                k = nCross, 
                n = nTrials,  
                grouping = c("vp_code","modality","v0","a","label","gain"),  
                fun=cum_normal_fun,  
                guess=0,  
                lapses=0,  
                bootstrap = 'none')  

qp_tidy <- tidyQuickPsy(qp)

qp_tidy$tidy_fit %>% 
  head()
```

tidyQuickPsy returns a list with two elements:  
-   `qp_tidy$qp` is the "old" quickpsy object
-   `qp_tidy$tidy_fit` is a tidy tibble containing the most important statistics of the quickpsy object.


### plotQuickPsy

Takes an object produced by tidyQuickPsy and plots the resulting psychometric functions for each person separately.

```{r, warning=FALSE, echo =TRUE}
qp_tidy <- tidyQuickPsy(qp)

plotQuickPsy(qp_tidy)$plot_list[[1]]

```

If the hessian matrix for one participant is not solvable, use this workaround to loop through all participant-condition combinations:

```{r, eval=FALSE, echo = TRUE}
# load data
data <- streetcrossing

# split by conditions and participant ID
data_list <- data %>% 
  group_split(vp_code,modality,v0,a,label,gain)

qp_list <- list()
df_list <- list()

# safe quickpsy execution
for(i in seq_along(data_list)){
  qp_list[[i]] <- tryCatch(tidyQuickPsy(quickpsy(d = data_list[[i]],
                           x = track_TTC,
                           k = nCross,
                           n = nTrials,
                           grouping = c("vp_code","modality","v0","a","label","gain"),
                           bootstrap = 'none')),
                           
                           warning = function(w){print(w)},
                           
                           error = function(e){e},
                           
                           finally = "finished")
  
  # tidy dfs
  df_list[[i]] <- qp_list[[i]]$tidy_fit
  
  status_message <- paste(round(i/length(data_list),4)*100,"% completed")
  print(status_message)
  if(rlang::is_error(qp_list[[i]])){
    print(paste("Error in data_list-element Nr.",i))
  }
}

# Filter List for errors
qp_list_clean <- qp_list[!sapply(qp_list, rlang::is_error)]
df_list_clean <- df_list[!sapply(df_list, rlang::is_error)]

# row bind data frame

df <- df_list_clean %>% 
  bind_rows()

# plot list

plot_list <- list()

for(i in seq_along(qp_list_clean)){
  plot_list[[i]] <- plotQuickPsy(qp_list_clean[[i]])$plot_list[[1]]

  status_message <- paste(round(i/length(qp_list_clean),2)*100,"% completed")
  print(status_message)
}
```
### plotThemeAGO

Plot theme for publication-ready data plots.

```{r, warning=FALSE, echo =TRUE}
ggplot(iris, aes(x = Petal.Length, y = Petal.Width, color = Species))+
  geom_point()+
  plotThemeAGO()
```


### Data sets


```{r, echo =TRUE}
data("loudness_block")
head(loudness_block)

data("streetcrossing")
head(streetcrossing)
```

